---
title: "BMKG - EDGAR"
author: "Centre for Research on Energy and Clean Air"
date: "10/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(root.dir=normalizePath('edgar/'))
```

# EDGAR Emission inventory

In this presentation, we'll download, analyse and plot EDGAR emission data in Indonesia. <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

## Load required libraries
```{r libraries, results="hide"}
library(readxl)
library(dplyr)
library(ggplot2)
library(tidyr)
library(reshape2)
if(!require(eixport)){devtools::install_github("atmoschem/eixport"); require(eixport)}

```


## Downloading data
We download data from [EDGAR](https://edgar.jrc.ec.europa.eu/overview.php?v=50_AP) using `eixport` package.

```{r download1, cache=T, message=F}
dir_edgar <- "/Volumes/ext1/data/emission/edgar/"

# Gridded data
eixport::get_edgar(pol=c("NOx","SO2","PM2.5"), year=2015, type="nc", destpath = dir_edgar, ask=F)
z <- list.files(dir_edgar, "*.zip", full.names = T)
f <- lapply(z, unzip, overwrite=F, exdir=dir_edgar)
file.remove(z)
```
```{r}
nc.paths <- list.files(dir_edgar, "*.nc", full.names = T)
nc.files <- reshape2::colsplit(gsub(paste(dir_edgar,".0.1x0.1.nc",sep="|"), "", ncs), "_", c("version","pollutant","year", "sector", "subsector")) %>%
  mutate(path=nc.paths,
         sector=ifelse(sector=="","Total",sector))

nc_to_province_emissions <- function(file, adm1){
  r <- raster::raster(file)
  s <- exact_extract(r, st_as_sf(adm1), 'weighted_sum', weights=area(r))
  s.kt.yr <- stat.nox * 3600 * 24 * 365
  result <- data.frame(adm1) %>% dplyr::select(province=NAME_1)
  result$emission.kt.yr <- s.kt.yr
  return(result)
}

province.emissions <- nc.files %>%
  rowwise() %>%
  mutate(province.emissions=list(nc_to_province_emissions(path, adm1))) %>%
  tidyr::unnest(province.emissions)

```

```{r}
ggplot(province.emissions %>% filter(sector=="Total")) +
  geom_bar(stat="identity", aes(x=emission.kt.yr, y=province, fill=sector)) +
  facet_wrap(~pollutant)

```

```{r download, cache=T, message=F}
# SO2
download.file("https://edgar.jrc.ec.europa.eu/download.php?edgar_dst=227347","so2.zip")
unzip(zipfile="so2.zip", files = "v50_SO2_1970_2015.xls", exdir=".")
d.so2 <- readxl::read_xls("v50_SO2_1970_2015.xls", skip = 8) %>%
    mutate(pollutant="so2") %>%
    filter(ISO_A3=="IDN")

# NO2
download.file("https://edgar.jrc.ec.europa.eu/download.php?edgar_dst=227343","nox.zip")
unzip(zipfile="nox.zip", files = "v50_NOx_1970_2015.xls", exdir=".")
d.nox <- readxl::read_xls("v50_NOx_1970_2015.xls", skip = 8) %>%
    mutate(pollutant="nox") %>%
    filter(ISO_A3=="IDN")

d.wide <- bind_rows(d.so2,
               d.nox)

d <- d.wide %>%
  tidyr::pivot_longer(cols=as.character(seq(1970, 2015)),
                      names_to="year") %>%
  mutate_at(c("year", "value"), as.numeric) %>%
  filter(!is.na(value))
```

## Plot timeseries in Indonesia


```{r plot1, echo=FALSE, fig.width = 10}
ggplot(d) +
  geom_area(aes(x=year, y=value, fill=IPCC_description)) +
  facet_wrap(~pollutant) +
  theme_light() +
  theme(legend.position="bottom") +
  labs(title="Emissions in Indonesia",
       y="kt/year",
       x=NULL,
       caption="Source: EDGAR v5.0")
```

```{r plot2, echo=FALSE, fig.width = 10, warning=FALSE}
ggplot(d %>% filter(year==2015) %>% arrange(pollutant, value)) +
  geom_bar(stat="identity", aes(x=value, y=reorder(IPCC_description, value))) +
  facet_wrap(~pollutant) +
  theme_light() +
  theme(legend.position="bottom") +
  labs(title="2015 Emissions in Indonesia",
       x="kt/year",
       y=NULL,
       caption="Source: EDGAR v5.0")
```

## Looking at provincial data
This time, we'll used the 2015 gridded data from EDGAR 5.0.

```{r, results="hide", message=F, warning=F}
library(raster)
library(rasterVis)
library(ncdf4)
library(GADMTools)
library(exactextractr)
adm1 <- GADMTools::gadm_sf_loadCountries("IDN", level=1, simplify=0.1)$sf %>% as_Spatial
```

### Maps
```{r, results="hide", message=F, warning=F}
download.file("https://edgar.jrc.ec.europa.eu/download.php?edgar_dst=226023","so2_grid.zip")
unzip(zipfile="so2_grid.zip", files = "v50_SO2_2015.0.1x0.1.nc", exdir=".")
r.so2 <- raster::raster("v50_SO2_2015.0.1x0.1.nc")
cuts <- c(.Machine$double.eps*1e13, 6e-4, 6e-2, 6e-1, 6, 30, 60, 240, 450, 1200)*1e-13 #set breaks
pal <- colorRampPalette(c("darkgreen","yellow", "red"))

plot(r.so2, breaks=cuts, col = pal(10), ext=bbox(adm1), box=F) + #plot with defined breaks
plot(adm1, add=T, lwd=0.6, border="#888888")
```

```{r, results="hide", message=F, warning=F}
download.file("https://edgar.jrc.ec.europa.eu/download.php?edgar_dst=217594","nox_grid.zip")
unzip(zipfile="nox_grid.zip", files = "v50_NOx_2015.0.1x0.1.nc", exdir=".")
r.nox <- raster::raster("v50_NOx_2015.0.1x0.1.nc")

cuts <- c(.Machine$double.eps*1e13, 6e-4, 6e-2, 6e-1, 6, 30, 60, 240, 450, 1200)*1e-13 #set breaks
pal <- colorRampPalette(c("darkgreen","yellow", "red"))

plot(raster::mask(r.nox, adm1), breaks=cuts, col = pal(10), ext=bbox(adm1), box=F) + #plot with defined breaks
plot(adm1, add=T, lwd=0.6, border="#888888")


```

```{r, results="hide", message=F, warning=F}
download.file("https://edgar.jrc.ec.europa.eu/download.php?edgar_dst=223979","pm25_grid.zip")
unzip(zipfile="pm25_grid.zip", files = "v50_PM2.5_2015.0.1x0.1.nc", exdir=".")
r.pm25 <- raster::raster("v50_PM2.5_2015.0.1x0.1.nc")

cuts <- c(.Machine$double.eps*1e13, 6e-4, 6e-2, 6e-1, 6, 30, 60, 240, 450, 1200)*1e-13 #set breaks
pal <- colorRampPalette(c("darkgreen","yellow", "red"))

plot(raster::mask(r.pm25, adm1), breaks=cuts, col = pal(10), ext=bbox(adm1), box=F) + #plot with defined breaks
plot(adm1, add=T, lwd=0.6, border="#888888")


```

### Statistics
```{r zonalstats}
# EDGAR data is given in g / m2 / s. Area in km2
# 1 g/m2/s * km2 = 1E3 kg / s =  t / s = (3600*24*365) t / yr
# We therefore weight every cell by its surface (not every pixel has the same surface)
stat.nox <- exact_extract(r.nox, st_as_sf(adm1), 'weighted_sum', weights=area(r.nox))
stat.not.kt.yr <- stat.nox * 3600 * 24 * 365

stat.so2 <- exact_extract(r.so2, st_as_sf(adm1), 'weighted_sum', weights=area(r.so2))
stat.so2.kt.yr <- stat.so2 * 3600 * 24 * 365

stat.pm25 <- exact_extract(r.pm25, st_as_sf(adm1), 'weighted_sum', weights=area(r.pm25))
stat.pm25.kt.yr <- stat.pm25 * 3600 * 24 * 365


stat = cbind(data.frame(adm1), nox=stat.not.kt.yr, so2=stat.so2) %>%
  tidyr::pivot_longer(cols=c(nox, so2), names_to="pollutant")
ggplot(stat) +
  geom_bar(aes(y=NAME_1, x=value), stat="identity") +
  facet_wrap(~pollutant) +
  labs(title="Total pollutant emission per province",
       y=NULL, x="kt / yr")
```


```{r}
stat.nox <- exact_extract(r.nox, st_as_sf(adm1), 'weighted_mean', weights=area(r.nox)) #kg/m2/s2
stat.nox.area <- exact_extract(area(r.nox), st_as_sf(adm1), 'sum') #km2
stat.nox.total <- stat.nox * stat.nox.area 
# 1 kg/m2/s * km2 = 1E6 kg / s = 1000 t / s = 1000*(3600*24*365) t / yr
stat.nox.total*1000*3600*24*365

stat.so2 <- exact_extract(r.so2, st_as_sf(adm1), 'weighted_sum', weights=area(r.nox))
stat.so2.area <- exact_extract(area(r.so2), st_as_sf(adm1), 'sum')
stat.so2.total <- stat.so2 * stat.so2.area



stat = cbind(data.frame(adm1), nox=stat.nox, so2=stat.so2) %>%
  tidyr::pivot_longer(cols=c(nox, so2), names_to="pollutant")
ggplot(stat) +
  geom_bar(aes(y=NAME_1, x=value), stat="identity") +
  facet_wrap(~pollutant) +
  labs(title="Sum of pollutant emission per province",
       y=NULL, x="kg / m2 / s")
```
